A Comparison of Graph Processing Systems

### 2020-09-16 documentation/paper/sections/results/galoisSpeedup.tex
hauptsächlich pr single node auswertung

\subsection{Galois speedup}
\label{sec:galois_speedup}
Analyzing the calculation time speedups for Galois, we can compare how or if the different algorithms benefit from increasing thread numbers.

\subsubsection{Single-source Shortest-path}
\begin{figure}
	\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisThreads.png}
	\caption{Calculation time speedup with increasing thread count for Galois Single-source Shortest-paths}
	\label{fig:galoisSpeedupSSSP}
\end{figure}
Starting with SSSP, we see an algorithm that benefits from many available threads in \autoref{fig:galoisSpeedupSSSP}.
For all larger graphs, speedup is in most cases very close to optimal up to about 8 threads. 
Twitter has the best speedup overall. It is 2.6$\times$ with 2 threads compared to one, 4$\times$ with 4, 7.7$\times$ with 8 and 9.7$\times$ using 16 threads.
Behaviour on friendster is similarly good. Here speedup is 1.9$\times$ at 2 threads compared to one, 3.5$\times$ at 4, 6.1$\times$ at 8 threads and 9.7$\times$ at 16 threads.
Anything above 16 threads however no longer helps decrease the computation time significantly on any graph. Speedup above 16 threads is always less than double the speedup of 16 threads. The maximum measured speedups are 10$\times$ (96 threads) for wikipedia, 17$\times$ (96 threads) for twitter, 11$\times$ (96 threads) for rMat27, 16$\times$ (48 threads) for friendster and 19$\times$ (40 threads) for rMat28.
In some cases increasing thread counts even prolongues calculation time. For example calculation on rMat28 is actually slower with 48 or 96 threads compared to 40 threads. For 40 threads, the speedup is nearly 19$\times$, on 48 threads 17$\times$ and with 96 threads only 15$\times$ compared to one thread.

@@

\subsubsection{Breadth-first search}
\begin{figure}
	\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisThreads.png}
	\caption{Calculation time speedup with increasing thread count for Galois Breadth-first search}
	\label{fig:galoisSpeedupBFS}
\end{figure}

@@

\subsubsection{PageRank}
We want to first take a look at the results for PageRank in Pull mode, seen in \autoref{fig:galoisSpeedupPRPull}. This is a perfect example for an algorithm that does not benefit from multithreaded computation. 
Computation is hardly sped up on any graph other than flicker, where the reached maximum is 64\%. This maximum is reached at two threads, with speedup steadily declining above that.
The rMat28 is the only other graph of one could say computation was sped up at large thread counts. Here we reached a maximum speedup of 31\%\ at 96 threads.
All 5 other graphs only reach a speedup greater or equal to 1 in just one or two cases and if so only by a small margin.
Computation on Orkut and Twitter reaches a speedup maximum of 12\%\ and 5\%\ at 4 threads, while being less or equal to 1 in all other cases.
The wikipedia graph is never sped up.
Friendster and rMat27 can be sped up by 6.5\%\ or 10\%\ respectively on 8 threads.

@@

\begin{figure}
	\begin{subfigure}{\columnwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisThreads.png}
		\caption{PageRank Pull}
		\label{fig:galoisSpeedupPRPull}
	\end{subfigure}
	\begin{subfigure}{\columnwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePRPushGaloisThreads.png}
		\caption{PageRank Push}
		\label{fig:galoisSpeedupPRPush}
	\end{subfigure}
	\caption{Calculation time speedup with increasing thread count for Galois PageRank Push and Pull algorithms.}
\end{figure}

\section{Hugepages}
\begin{figure}
	\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisHPThreads.png}
	\caption{Calculation time speedup with increasing thread count for Galois Single-source Shortest-paths with Hugepages}
	\label{fig:galoisHPSpeedupSSSP}
\end{figure}

\begin{figure}
	\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisHPThreads.png}
	\caption{Calculation time speedup with increasing thread count for Galois Breadth-first search with Hugepages}
	\label{fig:galoisHPSpeedupBFS}
\end{figure}

\begin{figure}
	\begin{subfigure}{\columnwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisHPThreads.png}
		\caption{PageRank Pull}
		\label{fig:galoisHPSpeedupPRPull}
	\end{subfigure}
	\begin{subfigure}{\columnwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePRPushGaloisHPThreads.png}
		\caption{PageRank Push}
		\label{fig:galoisHPSpeedupPRPush}
	\end{subfigure}
	\caption{Calculation time speedup with increasing thread count for Galois PageRank Push and Pull algorithms using Hugepages.}
\end{figure}

>>>>

\subsection{Galois speedup}
\label{sec:galois_speedup}
This section is dedicated to analyzing the speedup behaviour of Galois under the change to two parameters. First we change the thread count Galois is using. Second, a comparison between Galois using hugepages and Galois without hugepages is made.
We compare the \emph{calculation speedups} of Galois on the different graphs. Thus, we show the calculation time on any thread count normalized by the calculation time in the single-threaded environment.
Beginning with SSSP, followed by BFS and last the comparison for both PR Push and Pull.

\subsubsection{Single-Source Shortest-Paths}
\begin{figure*}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisThreads.png}
		\caption{without hugepages}
		\label{fig:galoisSpeedupSSSP_noHP}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisHPThreads.png}
		\caption{with hugepages}
		\label{fig:galoisSpeedupSSSP_HP}
	\end{subfigure}
	\hfil
	\caption{Calculation time speedups on SSSP for the different graphs with and without hugepages.}
	\label{fig:galoisSpeedupSSSP}
\end{figure*}
Starting with SSSP, we see an algorithm that benefits from many available threads (cf. \autoref{fig:galoisSpeedupSSSP}).
We first look at the speedups without hugepages, seen in \autoref{fig:galoisSpeedupSSSP_noHP}.
For all larger graphs, speedup is in most cases very close to optimal up to about 8 threads.
Twitter has the best speedup overall. It is 2.6$\times$ with 2 threads compared to one, 4$\times$ with 4, 7.7$\times$ with 8 and 9.7$\times$ using 16 threads.
Behaviour on friendster is similarly good. Here speedup is 1.9$\times$ at 2 threads compared to one, 3.5$\times$ at 4, 6.1$\times$ at 8 threads and 9.7$\times$ at 16 threads.
Anything above 16 threads however no longer helps decrease the computation time significantly on any graph. Speedup above 16 threads is always less than double the speedup of 16 threads. The maximum measured speedups are 10$\times$ (96 threads) for wikipedia, 17$\times$ (96 threads) for twitter, 11$\times$ (96 threads) for rMat27, 16$\times$ (48 threads) for friendster and 19$\times$ (40 threads) for rMat28.
In some cases increasing thread counts even prolongues calculation time. For example calculation on rMat28 is actually slower with 48 or 96 threads compared to 40 threads. For 40 threads, the speedup is nearly 19$\times$, on 48 threads 17$\times$ and with 96 threads only 15$\times$ compared to one thread.

@@

Upon activating the hugepages, we acquired the results seen in \autoref{fig:galoisSpeedupSSSP_HP}. Here, the overall results are similar to the findings without hugepages.

\todo{Hugepages}

\subsubsection{Breadth-First Search}
\begin{figure*}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisThreads.png}
		\caption{without hugepages}
		\label{fig:galoisSpeedupBFS_noHP}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisHPThreads.png}
		\caption{with hugepages}
		\label{fig:galoisSpeedupBFS_HP}
	\end{subfigure}
	\hfil
	\caption{Calculation time speedups on BFS for the different graphs with and without hugepages.}
	\label{fig:galoisSpeedupBFS}
\end{figure*}

@@

\subsubsection{PageRank}

We want to first take a look at the results for PageRank in Pull mode, seen in \autoref{fig:galoisSpeedupPRPull}. This is a perfect example for an algorithm that does not benefit from multithreaded computation.
Computation is hardly sped up on any graph other than flicker, where the reached maximum is 64\%. This maximum is reached at two threads, with speedup steadily declining above that.
The rMat28 is the only other graph of one could say computation was sped up at large thread counts. Here we reached a maximum speedup of 31\%\ at 96 threads.
All 5 other graphs only reach a speedup greater or equal to 1 in just one or two cases and if so only by a small margin.
Computation on Orkut and Twitter reaches a speedup maximum of 12\%\ and 5\%\ at 4 threads, while being less or equal to 1 in all other cases.
The wikipedia graph is never sped up.
Friendster and rMat27 can be sped up by 6.5\%\ or 10\%\ respectively on 8 threads.

@@

\todo{unfertig}
We assume the reason for this behaviour to be founded in the hugepages. Galois is recommended to be run with hugepages.

\begin{figure*}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisThreads.png}
		\caption{without hugepages}
		\label{fig:galoisSpeedupPRPull_noHP}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisHPThreads.png}
		\caption{with hugepages}
		\label{fig:galoisSpeedupPRPull_HP}
	\end{subfigure}
	\hfil
	\caption{Calculation time speedups on PR Pull for the different graphs with and without hugepages.}
	\label{fig:galoisSpeedupPRPull}
\end{figure*}

\begin{figure*}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePRPushGaloisThreads.png}
		\caption{without hugepages}
		\label{fig:galoisSpeedupPRPush_noHP}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePRPushGaloisHPThreads.png}
		\caption{with hugepages}
		\label{fig:galoisSpeedupPRPush_HP}
	\end{subfigure}
	\hfil
	\caption{Calculation time speedups on PR Pushgi for the different graphs with and without hugepages.}
	\label{fig:galoisSpeedupPRPush}
\end{figure*}

### 2020-09-16 documentation/paper/sections/results/pr.tex
kleine änderungen
hauptsächlich pr single node auswertung

\subsection{PageRank}
\subsubsection{Single-Node}

Ligra and Polymer support both regular and Delta-PageRank variants.
Ligra's regular PR implementation is faster on 4 of 7 graphs. If the regular version is slower than delta, that is only by a small difference. Explicitly, regular is slower than delta by 19\%\ on twitter, 17\%\ on rMat27 or 6\%\ for friendster. For the other graphs, the delta version is slower by a far greater margin, 50\%\ on flickr, 45\%\ on orkut, 13\%\ on wikipedia and 68\% on rMat28.
Hence, we only show the results of Ligra's regular PageRank implementation in our evaluation.
For Polymer we found the delta version to be faster on all graphs except rMat28. Delta-PR is on average 15\%\ faster on the first six graphs, while only being 0.3\%\ slower on rMat28.
Thus, the following only shows Polymer's faster Delta-PR implementation.

\todo{Hier fehlt noch Giraph, daher noch keine vollständige Auswertung.}

\begin{figure*}
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePR_calcTime.png}
		\caption{Calculation time}
		\label{fig:singleNodePR_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePR_execTime.png}
		\caption{Execution time}
		\label{fig:singleNodePR_exec}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePR_overheadTime.png}
		\caption{Overhead time}
		\label{fig:singleNodePR_overheadNormalized}
	\end{subfigure}

	\caption{Average times for PR on a single computation node, black bars represent one standard deviation in our testing}
\end{figure*}

\subsubsection{Distributed}
The \autoref{fig:distributedPR} shows our results of PageRank on the distributed cluster.
First of all, Giraph was unable to complete the test because it required more than 250GB of RAM for rMat28. Thus this results is missing.
When comparing the calculation times in \autoref{fig:distributedPR_calc} to the execution times in \autoref{fig:distributedPR_exec}, we see similar behaviour of all frameworks. This means that unlike with SSSP or BFS, the calculation times and execution times are similar with respect to the relations of the frameworks to one another. More specifically, there are no overhead outliers like it was the case with Giraph on SSSP.

\begin{figure*}
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedPR_calcTime.png}
		\caption{Calculation time}
		\label{fig:distributedPR_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedPR_execTime.png}
		\caption{Execution time}
		\label{fig:distributedPR_exec}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedPR_overheadTime.png}
		\caption{Overhead time}
		\label{fig:distributedPR_overheadNormalized}
	\end{subfigure}
	\caption{Average times for PR on the distributed cluster, black bars represent one standard deviation in our testing}
	\label{fig:distributedPR}
\end{figure*}

>>>>

\subsection{PageRank}

\subsubsection{Single-Node}
Ligra and Polymer support both regular and Delta-PageRank variants.
Ligra's regular PR implementation is faster on 4 of 7 graphs. If the regular version is slower than delta, that is only by a small difference. Explicitly, regular is slower than delta by a range of 6\% to 19\%\ on twitter, rMat27 or friendster. For the other graphs, the delta version is slower by a far greater margin of 13\% to 68\%.
Hence, we only show the results of Ligra's regular PageRank implementation in our evaluation.
For Polymer we found the delta version to be faster on all graphs except rMat28. Delta-PR is on average 15\%\ faster on the first six graphs, while only being 0.3\%\ slower on rMat28.
Thus, the following only shows Polymer's faster Delta-PR implementation.
Giraph required more than the available 250 GB of RAM on any graph larger than wikipedia, hence all of Giraph's results for the larger graphs are missing here.
\begin{figure*}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePR_calcTime.png}
		\caption{Calculation time}
		\label{fig:singleNodePR_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePR_execTime.png}
		\caption{Execution time}
		\label{fig:singleNodePR_exec}
	\end{subfigure}
	\hfil
	\caption{Average times for PR on a single computation node, black bars represent one standard deviation in our testing}
	\label{fig:singleNodePR}
\end{figure*}

The calculation times show some odd behaviour of Galois Push. The required time is less than 1ms, regardless of the graph (cf. \autoref{fig:singleNodePR_calc}). Meanwhile there was no output produced, that would indicate any kind of error. These results would make the calculation times of Galois Push the smallest on all graphs, with a difference of at least one order of magnitude. However, we are very suspicious of these results and thus exclude the calculation time for Galois Push in further comparisons. Because the execution time of Galois Push is always considerably longer than the execution time of Galois Pull (cf. \autoref{fig:singleNodePR_calc}). This leads us to believe that the output that we used for our measurements contains an error.

For the three graphs on which Giraph computed successfully, it is the slowest framework in both calculation and execution times. And that by a difference of three orders of magnitude in the calculation time and one to two orders of magnitude in execution time (cf \autoref{fig:singleNodePR}). On the larger graphs (i.e. those, where there is no data for Giraph), Gemini and Ligra are always slowest in execution time.
Contrary to this, Galois Pull has the smallest execution times on all graphs except flickr (cf. \autoref{fig:singleNodePR_exec}). Ligra is fastest on flickr, while being second fastest on wikipedia, rMat27 and rMat28. Galois Push is second fastest on orkut and friendster.
Interestingly, the execution time for Ligra is at a maximum for twitter. The required time is steadily decreasing with increasing graph size.

\subsubsection{Distributed}
The \autoref{fig:distributedPR} shows our results of PageRank on the distributed cluster.
First of all, Giraph was unable to complete the test because it required more than 250GB of RAM for rMat28, thus this result is missing.
When comparing the calculation times in \autoref{fig:distributedPR_calc} to the execution times in \autoref{fig:distributedPR_exec}, we see similar behaviour of all frameworks. This means that unlike with SSSP or BFS, the calculation times and execution times are similar with respect to the relations of the frameworks to one another. More specifically, there are no overhead outliers like it was the case with Giraph on SSSP.
\begin{figure*}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedPR_calcTime.png}
		\caption{Calculation time}
		\label{fig:distributedPR_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedPR_execTime.png}
		\caption{Execution time}
		\label{fig:distributedPR_exec}
	\end{subfigure}
	\hfil
	\caption{Average times for PR on the distributed cluster, black bars represent one standard deviation in our testing}
	\label{fig:distributedPR}
\end{figure*}

\todo{Auswertung fehlt noch.}
